# Dimentionality reduction techniques (week 5)



```{r}
date()
```

```{r}
library(dplyr)
library(ggplot2)
library(GGally)
library(corrplot)
library(tidyr)
library(ggpubr)
library(RColorBrewer)
library(FactoMineR)
library(factoextra)
```


## The graphical overview of the data

```{r}
setwd('//ad.helsinki.fi/home/z/zhangdon/Desktop/Introduction to open data science/IODS-project/data')
human <- read.table('human.csv', sep=',', header=TRUE, row.names = 1)

ggpairs(human)
cor(human) %>% corrplot()
```

There are 195 observations and 19 variables in the human data. According the instructions from last week, the column names of data had been shortened. For example, 
+ **gni_per** is gross national income per capita;
+ **le_birth** is life expectancy at birth;
+ **ey_edu** is expected years of schooling;
+ **mm_ratio** is maternal mortality ratio;
+ **ado_br** is adolescent birth rate;
+ **pp_parlia** is percentage of female representatives in parliament;
+ **edu2f** is proportion of females with at least secondary education;
+ **edu2m** is proportion of males with at least secondary education;
+ **labf** is proportion of females in the labor force;
+ **labm** is proportion of males in the labor force;
+ **edu2_ratio** is 'edu2f' / 'edu2m';
+ **lab_ratio** is 'labf' / 'labm'.

From ggpairs plot, most of the variables are normally distributed and significantly correlated with each other. The corrplot figure shows that maternal mortality ratio and adolescent birth rate have the negative correlations with most of variables. Red color represents the negative correlation while blue one means the positive correlation.

## PCA on the not scaled data

```{r, warning=FALSE, fig.height=15, fig.width=15}
pca_no_scaled <- prcomp(human)
s <- summary(pca_no_scaled)
s

pca_pr_no_scaled <- round(100*s$importance[2, ], digits = 1)
pc_lab_no_scaled <- paste0(names(pca_pr_no_scaled), " (", pca_pr_no_scaled, "%)")
biplot(pca_no_scaled, , cex = c(0.8, 1), col = c("grey40", "deeppink2"), xlab = pc_lab_no_scaled[1], ylab = pc_lab_no_scaled[2])
```

## PCA on the scaled data

```{r, warning=FALSE, fig.height=15, fig.width=15}
human_std <- scale(human)
summary(human_std)

pca_scaled <- prcomp(human_std)
s1 <- summary(pca_scaled)
s1

pca_pr_scaled <- round(100*s1$importance[2, ], digits = 1)
pc_lab_scaled <- paste0(names(pca_pr_scaled), " (", pca_pr_scaled, "%)")
biplot(pca_scaled, , cex = c(0.8, 1), col = c("grey40", "deeppink2"), xlab = pc_lab_scaled[1], ylab = pc_lab_scaled[2])
```

Compared with plots with and without standardization, the scaled data show the reasonable consequence about the correlation among variables. In the not-scaled figure, **GNI** is the main factor for PCA, the priciple componet 1 accounts for 100% of all principle components. However, in the scaled figure, PC1 is 53.6%, pc2 is 16.2%. All scaled variables seem reasonable. **pp_parlia** and **lab_ratio** are mostly contributed to the PC2 and have the higher positive correlations, the female ratio is the main point in these two variables. Others are related to PC1. **edu2_ratio**, **ey_edu**, **GNI** and **le_birth** are correlated with each other, whereas **mm_ratio** and **ado_br** are closer because the death of mother would cause the birth of new infants. 

## Principal component dimensions

```{r}
s # not scaled
s1 # scaled
```

According to the importance of components, the standard deviation and proportion of variance for principle components are the highest. From the biplot figure, 6 of variables are correlated to the PC1 because the angles between are relatively smaller. The length of **GNI** and **edu2_ratio** arrows are shorter than others, which means that their standard deviations are smaller than others accordingly.

## MCA on the tea dataset

```{r, warning=FALSE}
data(tea)
keep_columns <- c("Tea", "How", "how", "sugar", "where", "lunch")
tea_time <- select(tea, one_of(keep_columns))
str(tea_time)
dim(tea_time)
gather(tea_time) %>% ggplot(aes(value)) + geom_bar() + facet_wrap("key", scales = "free") + theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))
```

There are 300 observations and 6 variables in **tea_time** data. All of variables are the factor type data which are categorized into 2-4 levels. From the ggplot figure, tea bag is the popular way for tea drinking. Most of people like drinking tea without any additions and not with lunch. There is no difference from sugar and no sugar, however, the most favorite tea and drinking area are Earl Grey and chain store, respectively.


```{r}
mca <- MCA(tea_time, graph = FALSE)
summary(mca)
```

After Multiple Correspondence Analysis, The eigenvalues, individuals and categories of each dimensions can be summarized.

```{r, warning=FALSE}
fviz_screeplot(mca, addlabels = TRUE, ylim = c(0, 45))
var <- get_mca_var(mca)
```

15.2% variances are accounting for dimension1, then 14.2 for dimension2 and 12% for dimension3. The change of explained variances in each dimensions are not dropped sharply, the differences among them are slight. Thus, it is necessary to explore the details of different dimensions.

```{r, warning=FALSE, fig.height=9, fig.width=9}
fig1 <- fviz_contrib(mca, choice = "var", axes = 1, top = 15)
fig2 <- fviz_contrib(mca, choice = "var", axes = 2, top = 15)
fig3 <- fviz_contrib(mca, choice = "var", axes = 3, top = 15)
fig4 <- fviz_contrib(mca, choice = "var", axes = 4, top = 15)

ggarrange(fig1, fig2, fig3, fig4,
          ncol = 2, nrow = 2)
```

In dimension1, **tea shop** and **unpacked tea** are the main contributors. In dimension2, **chain store + tea shop** and **tea bag + unpackaged** are the most important factors. In dimension3 and 4, type of tea is the significant contributor, black tea and green tea have the highest contributions in these two dimensions separately.

```{r, warning=FALSE}
fviz_contrib(mca, choice = "var", axes = 1:4, top = 15)
```

Among the first 4 dimensions,  **tea shop**, **unpacked tea**, **chain store + tea shop**, **tea bag + unpackaged** have the higher contribution.

```{r, warning=FALSE}

fviz_mca_var(mca, col.var = "cos2",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             repel = TRUE,
             ggtheme = theme_minimal())

```

According to *cos2* values, when it is closer to 1 means the strong correlations between the variable and dimension. On the above figure, **chain store + tea shop** and **tea bag + unpackaged**, **tea bag** and **chain store**, **unpackaged** and **tea shop** are the most quality variables to dimensions.

```{r, warning=FALSE, fig.height=8, fig.width=8}
plot(mca, invisible=c("ind"), habillage = "quali", cex = 1.3, palette=palette(brewer.pal(8, 'Dark2')))
```

Definitely, two groups (**how** - the pack of tea and **where** - the drinking area) are much similar with each other.
